Planetary perturbers: Flaring star-planet interactions in Kepler and TESS

### 2023-10-12 src/tex/ms.tex
[showyourwork] overleaf sync

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view. For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
Equivalent duration is equal to the time during which the non-flaring star releases as much energy as the flare.

@@

\begin{equation}
    P_{\rm spi,sb} \propto B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3} R_*^{(n+11) / 3} v_{\rm rel} a^{-(n+11) / 3},
\end{equation}

\textcolor{red}{ignoring factors of the order of 1.}  $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=XXXX$, which best reproduces observations of XXXX in~\citet{lanza2012starplanet}. Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity.} We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Section~\ref{sec:methods:bfield}) as a proxy.

@@

\begin{equation}
    P_{\rm spi,sb0} \propto R_{\rm p}^2 R_*^{(n+11) / 3} B_*^2 v_{\rm rel}    a^{-(n+11) / 3} 
\end{equation}

@@

\begin{equation}
    P_{\rm spi,aw} \propto B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta
\end{equation}

\textcolor{red}{Here, we assume that the magnetopause radius extends to where the magnetic pressures of planet and stellar magnetic wind are equal, and the wind is dominated by the magnetic field in de sub-Alf\'enic regime. We approximate the stellar wind density $\rho_W$ with a proportionality to $\rho_* (R_* / a)^{2}$, where $\rho_*$ is the stellar wind density at the base of the corona. For the stellar wind magnetic field at the planetary radius $B_W$, we insert $B_* (R_* / a)^{3}$, i.e., the field strength decreasing radially outward from the surface as a dipole. This produces an upper bound on the field strength at the planet's distance because more complex fields would drop off more steeply. The last two replacements result in the total $R_*^2 a^{-2}$ factor in the above equation. Since $\rho_*$ is poorly constrained empirically~\citep{vidotto2021evolution}, we adopt $\rho_*=2\cdot10^{10}\rm{cm}^{-3}$, a value often used for the Sun~(XXXX), and for stellar simulations~\citep{alvarado-gomez2020tuning,kavanagh2021planetinduced}. The angle between the wind magnetic field vector and the velocity vector of the planet $\theta$ is also unknown, so we set $\sin\theta=1$ to obtain an upper limit. In reality, $\theta$ will vary both between systems, and over time in any individual system. Omitting $\rho_*$ assumes that all stellar wind densities are the same in our sample, which is not true in reality. The dependence on $\rho_*$ is stronger than that on $B_*$, but weaker than on every other parameter, so it may dominate the magnetic field dependence.}

@@

Figures~\ref{fig:adtest_bp} and \ref{fig:adtest_bp_rest} show that in every scenario we considered (Alfv\'en wing or stretch-and-break mechanism, with magnetized or unmagnetized planets, see Section~\ref{sec:methods:pspi}), there are two branches on the high end of expected powers. On one branch, the significance of the measured interaction increases with the expected power $P_{\rm spi}$. On the other branch, no interaction is measured regardless of $P_{\rm spi}$. \textcolor{red}{For all powers of SPI, we normalize the values to AU Mic, and give an overview over the normalization values in Table~\ref{tab:aumic}}.

@@

\section{Flaring finding in Kepler and TESS: False positives}

\begin{figure*}
    \script{paper_false_positives.py}
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/systematics_false_positives.png}
    \caption{\textcolor{red}{True and false positive flare candidates in Kepler and TESS light curves. \textbf{Left panel}: A flare in the TESS 2-min cadence light curve of GJ 3323 in Sector 5, with its characteristic fast-rise-exponential-decay shape.  \textbf{Middle panel}: Kepler-235 with a possible argabrightening in the 1-min cadence light curve in  Quarter 14. Argabrightenings are few-minutes long reflections of light off debris and onto the detector~\citep{vancleve2016kepler}. The reflections affect large parts of the detector, so that this event can be identified as a false positive by its common occurrence across many light curves. \textbf{Right panel:} K2-77 with a Solar System Object (SSO) moving across the TESS detector in the 2-min cadence light curve in Sector 44. SSOs produce smooth and nearly symmetrical brightenings as they move in and out of the aperture of the star. They can also (partially) be identified with known SSOs~\citep{berthier2006skybot, berthier2016prediction}.}}
    \label{fig:flares_and_fps}
        \end{centering}
\end{figure*}

>>>>

We vet all flare candidates by eye, and exclude instrumental and physical false positives, such as Solar System Objects passing by the field of view~(see Fig.~\ref{fig:flares_and_fps} in the Appendix for an example). For each confirmed flare candidate, we calculate the relative amplitude $a$ and equivalent duration $ED$, defined as the integrated flare flux $F_{\rm flare}$ over the duration of the flare, divided by the median flux $F_0$ of the star, integrated over the flare duration~\citep{gershberg1972results}:
\begin{equation}
\label{eq:ED}
ED=\displaystyle \int \mathrm dt\, \frac{F_{\rm flare}(t)}{F_0}.
\end{equation}
Equivalent duration is equal to the time during which the non-flaring star releases as much energy as the flare.

@@

\begin{equation}
    P_{\rm spi,sb}=\underbrace{\frac{27\pi \lambda^2(n+1)}{16 \mu_0 n (\lambda^2 + n^2)^{1/3}}}_{\hat{C}} B_{\rm p}^{2/3} R_{\rm p}^2 B_*^{4/3}  v_{\rm rel} \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}},
\end{equation}

\textcolor{red}{where $\mu_0$ is the magnetic permeability of the vacuum, and $n$ and $\lambda$ are dimensionless constants, all of which we subsume under $\hat{C}$ together with other constant factors}. $B_{\rm p}$ and $B_*$ are the planetary and stellar field strengths at their poles, respectively; $v_{\rm rel}$ is the relative velocity between the stellar rotation and the planet's orbit at the semi-major axis $a$; and $R_*$ and $R_{\rm p}$ are the stellar and planetary radii, respectively. \textcolor{red}{We choose $n=0.25$ ($\lambda^2 = 1.01203$, which best reproduces observations of SPI candidate systems HD~179949, $\tau$ Boo, and HD~189733 in~\citet{lanza2012starplanet} (except for $\upsilon$ And). Since the planetary fields are unknown, we adopt $B_{\rm p}=1\,$G for simplicity.} We also do not know the stellar field strength at the pole for the vast majority of the stars, so we adopt \textcolor{red}{an average value of 15 per cent of the} surface field strength~(see Section~\ref{sec:methods:bfield}) as a proxy.

@@

\begin{equation}
    P_{\rm spi,sb0} =\hat{C} R_{\rm p}^2 B_*^2 v_{\rm rel}  \left(\frac{R_*}{a}\right)^{\frac{n+11}{ 3}}
\end{equation}

@@

\begin{equation}
    P_{\rm spi,aw} = \pi^{1/2} B_{\rm p}^{2/3} R_{p}^2  R_*^2  B_*^{1/3}  v_{\rm rel}^2 a^{-2} \rho_*^{1/2} \sin^2 \theta
\end{equation}

\textcolor{red}{Here, we assume that the magnetopause radius extends to where the magnetic pressures of planet and stellar magnetic wind are equal, and the wind is dominated by the magnetic field in de sub-Alf\'enic regime. We approximate the stellar wind density $\rho_W$ with a proportionality to $\rho_* (R_* / a)^{2}$, where $\rho_*$ is the stellar wind density at the base of the corona. For the stellar wind magnetic field at the planetary radius $B_W$, we insert $B_* (R_* / a)^{3}$, i.e., the field strength decreasing radially outward from the surface as a dipole. This produces an upper bound on the field strength at the planet's distance because more complex fields would drop off more steeply. The last two replacements result in the total $R_*^2 a^{-2}$ factor in the above equation. Since $\rho_*$ is poorly constrained empirically~\citep{vidotto2021evolution}, we adopt $\rho_*=2\cdot10^{10}\rm{cm}^{-3}$, a value often used for the Sun~\citep{sokolov2013magnetohydrodynamic}, and for stellar simulations~\citep{alvarado-gomez2020tuning,kavanagh2021planetinduced}. The angle between the wind magnetic field vector and the velocity vector of the planet $\theta$ is also unknown, so we set $\sin\theta=1$ to obtain an upper limit. In reality, $\theta$ will vary both between systems, and over time in any individual system. Omitting $\rho_*$ assumes that all stellar wind densities are the same in our sample, which is not true in reality. The dependence on $\rho_*$ is stronger than that on $B_*$, but weaker than on every other parameter, so it may dominate the magnetic field dependence.}

@@

Figures~\ref{fig:adtest_bp} and \ref{fig:adtest_bp_rest} show that in every scenario we considered (Alfv\'en wing or stretch-and-break mechanism, with magnetized or unmagnetized planets, see Section~\ref{sec:methods:pspi}), there are two branches on the high end of expected powers. On one branch, the significance of the measured interaction increases with the expected power $P_{\rm spi}$. On the other branch, no interaction is measured regardless of $P_{\rm spi}$.

@@

\newpage
\section{Flaring finding in Kepler and TESS: False positives}
\restartappendixnumbering

\begin{figure*}[ht!]
    \script{paper_false_positives.py}
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/systematics_false_positives.png}
    \caption{\textcolor{red}{True and false positive flare candidates in Kepler and TESS light curves. \textbf{Left panel}: A flare in the TESS 2-min cadence light curve of GJ 3323 in Sector 5, with its characteristic fast-rise-exponential-decay shape, with additional substructure around the peak. \textbf{Middle panel}: Kepler-235 with a possible argabrightening in the 1-min cadence light curve in  Quarter 14. Argabrightenings are few-minutes long reflections of light off debris and onto the detector~\citep{vancleve2016kepler}. The reflections affect large parts of the detector, so that this event can be identified as a false positive by its common occurrence across many light curves. \textbf{Right panel:} K2-77 with a Solar System Object (SSO) moving across the TESS detector in the 2-min cadence light curve in Sector 44. SSOs produce smooth and nearly symmetrical brightenings as they move in and out of the aperture of the star. They can also (partially) be identified with known SSOs~\citep{berthier2006skybot, berthier2016prediction}.}}
    \label{fig:flares_and_fps}
        \end{centering}
\end{figure*}
